Make any stupid plot you want

### 2021-02-25 draft/sections/implementation.tex
nominally done (no figure) but won't compile

The \mintinline{python}{Point} artist builds on \mintinline{python}{collection} artists because collections are optimized to efficiently draw a sequence of primitive point and area marks. In this prototype, the scatter marker shape is fixed as a circle, and the only visual fiber components are x and y position, size, and the facecolor of the marker. We only show the \mintinline{python}{assembly} function here because the \mintinline{python}{__init__, draw} are identical the prototype artist shown in \ref{code:artist}. 
\begin{minted}{python}
class Point(mcollections.Collection):
    def assemble(self, x, y, s, facecolors='C0' ):
        # construct geometries of the circle glyphs in visual coordinates
        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                    for (xi, yi, si) in zip(x, y, s)] 
        # set attributes of glyphs, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(facecolors)
\end{minted} 
The \mintinline{python}{view} method repackages the data as a fiber component indexed table of vertices, as described in section~\ref{sec:triangulization}. Even though the \mintinline{python}{view} is fiber indexed, each vertex at an index \dbasepoint has corresponding values in section $\dsection(\dbasepoint_{i})$. This means that all the data on one vertex maps to one glyph. To ensure the integrity of the section, \mintinline{python}{view} must be atomic. This means that the values cannot change after the method is called in draw until a new call in draw. We put this constraint on the return of the \mintedinline{python}{view} method so that we do not risk race conditions.

@@

\begin{minted}{python}

\end{minted}
The \mintinline{python}{draw} method here has a more complex unpacking of visual encodings to support passing in visual component data directly. This is necessary to build composite artists out ouf of this bar artists.

This bar underpins a rudimentary version of the complex \plus operator such that we can specify multiple components to map to the same visual components without requiring th components to be individually specified in the \mintinline{python}{transforms} dictionary. The implementation of this support is slightly more complicated than the simple case, but not in a way that is useful for understanding the framework. The \mintinline{python}{assemble} function constructs bars and sets their edge color to black. Defaults are provided for 'width' and 'floor' to make this function more reusable. Typically the defaults are used for the type of chart shown in figure~\ref{fig:code_bar_simple}, but these visual variables are often set when building composite versions of this chart type as discussed in section~\ref{sec:code_case_study}.

\subsection{Encoders \vchannel}
\label{sec:code_channels}
As mentioned above, the encoding dictionary is specified by the visual fiber component, the corresponding data fiber component, and the mapping function. The visual parameter serves as the dictionary key because the visual representation is constructed from the encoding applied to the data  $\vsection = \vchannel \circ \dsection$. For the scatter plot, the mappings for the visual fiber components $\vfiber=(x,y, facecolors, s)$ are defined as
\begin{minted}{python}
cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
transforms = {'y': ('v1', lambda x: x),       
              'x': ('v3', lambda x: x),
             'facecolors': ('v2', cmap), 
             's':(None ,lambda _: itertools.repeat(.02))}
\end{minted}
where the position \textit{(x,y)} \vchannel\ transformers are identity functions. The size \textit{s} transformer is not acting on a component of \dfiber, instead it is a \vchannel\ that returns a constant value. While size could be embedded inside the \mintinline{python}{assembly} function, it is added to the transformers to illustrate user configured visual parameters that could either be constant or mapped to a component in \dfiber. The identity and constant \vchannel\ are explicitly implemented here to demonstrate their implicit role in the visual pipeline, but they could be optimized away. More complex encoders can be implemented as callable classes, such as

\begin{minted}{python}
class Categorical:
    def __init__(self, mapping):
        # check that the conversion is to valid colors
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

@@

where \mintinline{python}{__init__} can validate that the output of the \vchannel\ is a valid element of the \vfiber\ component the \vchannel\ function is targeting. Creating a callable class also provides a simple way to swap out the specific (data, value) mapping without having to reimplement the validation or conversion logic. A test for equivariance can be implemented trivially
\begin{minted}{python}
def test_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    assert sorted(m1) == sorted(m2)
\end{minted}
but is currently factored out of the artist for clarity. In this example, \mintinline{python}{is_nominal} checks for equivariance of permutation group actions by applying the encoder to a set of values, shuffling values, and checking that (value, encoding) pairs remain the same. This equivariance test can be implemented as part of the artist or encoder, but for minimal overhead, the equivariant it is implemented as part of the library tests.

@@

    def view(self, axes):
        table = defaultdict(list)
        for k in self.keys:
            table['index'] = k
            # on each iteration, add one (name, value) pair per component
            for (name, value) in zip(self.FB.fiber.keys(), self.tau(k)[1]):
                table[name].append(value)
        return table
\end{minted}
The view method returns a dictionary where the key is a fiber component name and the value is a list of values in the fiber component. The table is built one call to \mintinline{python}{tau} at a time, guaranteeing that all the fiber component values are over the same \dbasepoint. Table has a \mintinline{python}{get} method as it is a method on Python dictionaries. In contrast, the line in \mintinline{python}{EdgeSimplex} is defined as the functions \mintinline{python}{_color,_xy} on each edge.

@@

    def view(self, axes):
        table = defaultdict(list)
        for k in self.keys:
            table['index'].append(k)
            # (name, value) pair, value is [x0, ..., xn] for x, y
            for (name, value) in zip(self.FB.fiber.keys(), self.tau(k, simplex)[1]):
                table[name].append(value)
\end{minted}
Unlike scatter, the line \mintinline{python}{tau} method returns the functions on the edge evaluated on the interval [0,1]. By default these means each \mintinline{python}{tau} returns a list of 1000 x and y points and the associated color. As with scatter, \mintinline{python}{view} builds a table by calling \mintinline{python}{tau} for each \dbasepoint\. Unlike scatter, the line table is a list where each item contains a list of points. This bookkeeping of which data is on an edge is used by the \mintinline{python}{assembly} functions to bind segments to their visual properties.

@@

\subsection{Case Study: Penguins}
For this case study, we use the Palmer Penguins dataset\cite{gormanEcologicalSexualDimorphism2014, horstPalmerpenguinsPalmerArchipelago2020} since it is multivariate and has a varying number of penguins. We use a version of the data packaged as a pandas dataframe\cite{nakhaeeMcnakhaeePalmerpenguins2021, jeff_reback_2020_3715232} since that is a very commonly used Python labled data structure. The wrapper is very thin since here there is explicitly only one section.
\begin{minted}{python}
class DataFrameSection:
    def __init__(self, dataframe):
        self._tau = dataframe.iloc
        self._view = dataframe
    def view(self, axes):
        return self._view
\end{minted}
The pandas indexer is a key valued set of discrete vertices, so there is no need to repackage for triangulation.

@@

For example, the \mintinline{python}{MultiBar} artist that makes figure~\ref{fig:code_bar_multi} reuses \mintinline{python}{_make_bars} but does not reuse the assemble function because the composition of elements forces fundemental differences in glyph construction. As demonstrated in the \mintinline{python}{init}, the composite bar chart has orientation and whether it is stacked or not. While the stacked bar chart and the grouped bar chart could be seperate artists, as demonstrated they share so much overlapping code that it is far less redundant to implement them together.  \note{looking at the mess that is this code, I'm a) not convinced these should be combined b) no longer convinced this provides anything over just bar if it isn't rewritten to use bar more}

\begin{minted}{python}
class MultiBar(mcollections.Collection):
    def __init__(self, data, transforms, *args, **kwargs):
        #set the orientation of the graphic  
        self.orientation = kwargs.pop('orientation', 'v')
        # set how the bar glyphs are put together to create the graphic
        self.stacked = kwargs.pop('stacked', False)     
        # rest is same as other artist __init__s

        #this needs to be factored out but just want to finish now
        self.width = kwargs.pop('width', .8)

    def assemble(self, visual, view):
        (groups, gencoder) = self.transforms['length']
        ngroups = len(np.atleast_1d(groups))
        visual['floor'] = visual.get('floor', np.empty(len(view[groups[0]])))
        visual['facecolors'] = visual.get('facecolors', 'C0')
        # make equal width stacked columns 
        if 'width' not in visual and self.stacked:
            visual['width'] = itertools.repeat(self.width)

        # make equal with groups
        if not self.stacked:
            visual['width'] = itertools.repeat(self.width/ngroups)
            offset = (np.arange(ngroups) /ngroups) * self.width
        else:
            offset = itertools.repeat(0)

        # make the bars and arrange them 
        verts = []
        for group, off in zip(groups, offset):
            verts.extend(Bar._make_bars(self.orientation, visual['position'] + off, 
                          visual['width'], visual['floor'], view[group]))
            if self.stacked: #add stacked bar to previous bar
                visual['floor'] += view[group]

        # convert lengths after all calculations are made and reorient if needed
        # here or in transform machinery?
        if self.orientation in {'v', 'vertical'}:
            tverts = [[(x, gencoder(y)) for (x, y) in vert] 
                            for vert in verts]
        elif self.orientation in {'h', 'horizontal'}:
            tverts = [[(gencoder(x), y) for (x, y) in vert] 
                            for vert in verts]
        self._paths = [mpath.Path(xy, closed=True) for xy in tverts]
        #flatted columns of colors to match list of bars
        self.set_facecolor(list(itertools.chain.from_iterable(visual['facecolors'])))
        self.set_edgecolors('k')

    def draw(self, renderer, *args, **kwargs):
        view = self.data.view()
        #exclude converting the group visual length, special cased in assemble
        visual = utils.convert_transforms(view, self.transforms, exclude=['length'])
        # pass in view because nu is not distributable so may need to apply it 
        # after visual assembly
        self.assemble(visual, view)
        super().draw(renderer, *args, **kwargs)
        return      
\end{minted}

In the \mintinline{python}{__draw__}, a utility function is used for conversions, but the length transforms are held until after assembly because the length is computed by adding the current length to the previous and many transforms are not distributable such that $\nu(x_0 + x_1 + x_2) = \nu(x_0) + \nu(x_1) + \nu(x_2)$. Inside \mintinline{python}{assemble}, the glyphs are either shifted vertically (\texttt{stacked}) or horizontally (\texttt{grouped}) such that the positions are recorded and added to with the next group. This function allows multiple columns to be mapped to a visual parameter, but it must be equal numbers of columns
\begin{minted}{python}
{'position': ('island', lambda x: {'Biscoe':0, 'Dream':1, 'Torgersen':2}[x]),
 'length':(['Adelie', 'Chinstrap', 'Gentoo'], lambda x: x),
 'facecolors': (['Adelie_s', 'Chinstrap_s', 'Gentoo_s'], 
        color.Categorical({'Adelie':'#FF8C00',
                           'Gentoo':'#159090', 
                           'Chinstrap':'#A034F0'}))}
\end{minted}
such as in this example where for each column contributing to a segment of the bar there is a corresponding column of colors for this segment. The reason the multibar can work with such a trasnformer is because it is relying on the data model to do most of the bookkeeping of which values get mapped to which bars. This also yields a much simpler function call to the artist

\begin{minted}{python}
fig, ax = plt.subplots()
artist = bar.MultiBar(table, trans, orientation='h', stacked=True)
ax.add_artist(artist)
\end{minted}

where \mintinline{python}{trans} is the same dictionary for both stacked and grouped version, as is the \mintinline{python}{DataFrameSection} object \mintinline{python}{table}. The only difference between the two versions is the \texttt{stacked} flag, and the only difference between figures~\ref{fig:code_bar_simple} is the \texttt{orientation} argument. By decomposing the architecture into data, visual encoding, and assembly steps, we are able to build components that are more flexible and also more self contained than the existing code base.

\subsection{Summary}
In general, the way in which we implemented the artist is as follows: 
\begin{description}
    \item[\textbf{\vchannel}] encoder function converting data to library normalized form, can also be \mintinline{python}{Encoder.__call__} method
    \item[\textbf{\vsection}]  dictionary of the form \mintinline{python}|{parameter:(variable, encoder)}| curried until {Artist.draw}
    \item[\textbf{\vartisteq}] objects that take in as input \dtotal\ and \vtotal\, for example \mintinline{python}{Point}, \mintinline{python}{Line}, \mintinline{python}{bar}
    \item[\textbf{\vartist}] \mintinline{python}{Artist.draw} method which is \vartisteq\ parameterized by \vtotal
    \item[\textbf{\vmarkd}] \mintinline{python}{Artist.assemble(visual)} method that arranges components of \vtotal\ into glyphs  
\end{description}

with the data representation implemented as 
\begin{description}
    \item[\textbf{\dtotal}] object with \mintinline{python}{view} method 
    \item[\textbf{\dfiber}] named components of the data and their types
    \item[\textbf{\dbase}] determines how the data is bound together when returned by view  
    \item[\textbf{\dsection}] \mintinline{python}{Data.view()} method that returns component sections
\end{description}

While very rough, this API demonstrates that the ideas presented in the math framework are implementable. In choosing a functional approach, if not implementation, we provide a framework for library developers to build reusable encoder \vchannel\ and assembly \vmarkd\. We argue that  if these functions are built such that they are equivariant with respect to monoid actions and the graphic topology is a deformation retraction of the data topology, then the artist by definition will be a structure and property preserving map from data to graphic. 
\end{document}

>>>>

The \mintinline{python}{Point} artist builds on \mintinline{python}{collection} artists because collections are optimized to efficiently draw a sequence of primitive point and area marks. In this prototype, the scatter marker shape is fixed as a circle, and the only visual fiber components are x and y position, size, and the facecolor of the marker. We only show the \mintinline{python}{assembly} function here because the \mintinline{python}{__init__, draw} are identical the prototype artist shown in \ref{code:artist}. 
\begin{minted}{python}
class Point(mcollections.Collection):
    def assemble(self, x, y, s, facecolors='C0' ):
        # construct geometries of the circle glyphs in visual coordinates
        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                    for (xi, yi, si) in zip(x, y, s)] 
        # set attributes of glyphs, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(facecolors)
\end{minted} 
The \mintinline{python}{view} method repackages the data as a fiber component indexed table of vertices, as described in section~\ref{sec:triangulization}. Even though the \mintinline{python}{view} is fiber indexed, each vertex at an index \dbasepoint has corresponding values in section $\dsection(\dbasepoint_{i})$. This means that all the data on one vertex maps to one glyph. To ensure the integrity of the section, \mintinline{python}{view} must be atomic. This means that the values cannot change after the method is called in draw until a new call in draw. We put this constraint on the return of the \mintinline{python}{view} method so that we do not risk race conditions.

@@

\begin{minted}{python}
class Bar(mcollections.Collection):
    def __init__(self, data, transforms, orientation='v', *args, **kwargs):
        """
        orientation: str, optional
            v: bars aligned along x axis, heights on y
            h: bars aligned along y axis, heights on x
        """
        self.orientation = orientation
        super().__init__(*args, **kwargs)
        self.data = data
        self.transforms = copy.deepcopy(transforms)

    def assemble(self, position, length, floor=0, width=0.8, facecolors='C0', edgecolors='k', offset=0):
        #set some defaults
        width = itertools.repeat(width) if np.isscalar(width) else width
        floor = itertools.repeat(floor) if np.isscalar(floor) else (floor)

        # offset is passed through via assemblers such as multigroup, not supposed to be directly tagged to position 
        position = position + offset

        def make_bars(xval, xoff, yval, yoff):
             return [[(x, y), (x, y+yo), (x+xo, y+yo), (x+xo, y), (x, y)] 
                for (x, xo, y, yo) in zip(xval, xoff, yval, yoff)]
        #build bar glyphs based on graphic parameter
        if self.orientation in {'vertical', 'v'}:
            verts = make_bars(position, width, floor, length)
        elif self.orientation in {'horizontal', 'h'}:
            verts = make_bars(floor, length, position, width)

        self._paths = [mpath.Path(xy, closed=True) for xy in verts]
        self.set_edgecolors(edgecolors)
        self.set_facecolors(facecolors)

    def draw(self, renderer,  *args, **kwargs):
        view = self.data.view(self.axes)
        visual = {}
        for (p, t) in self.transforms.items():
            if isinstance(t, dict):
                if t['name'] in self.data.FB.F and 'encoder' in t:
                    visual[p] = t['encoder'](view[t['name']])
                elif 'encoder' in t: # constant value
                    visual[p] = t['encoder'](t['name'])
                elif t['name'] in self.data.FB.F: # identity 
                    visual[p] = view[t['name']]
            else: # no transform 
                 visual[p] = t
        self.assemble(**visual)
        super().draw(renderer,  *args, **kwargs)

\end{minted}
The \mintinline{python}{draw} method here has a more complex unpacking of visual encodings to support passing in visual component data directly. This is vastly simplifies building composite objects as the alternative would be higher order functions that take as input the transforms passed in by the user. This construction supports a constant visual parameter, an identity transform where the value is the same in \dtotal\ and \vtotal\, and setting the visual component directly. The \mintinline{python}{assemble} function constructs bars and sets their face and edge colors. The \mintinline{python}{make_bars} function converts the input position and length to the coordinates of a rectangle of the given width. Defaults are provided for 'width' and 'floor' to make this function more reusable. Typically the defaults are used for the type of chart shown in figure~\ref{fig:code_bar_simple}, but these visual variables are often set when building composite versions of this chart type as discussed in section~\ref{sec:code_case_study}.

\subsection{Encoders \vchannel}
\label{sec:code_channels}
As mentioned above, the encoding dictionary is specified by the visual fiber component, the corresponding data fiber component, and the mapping function. The visual parameter serves as the dictionary key because the visual representation is constructed from the encoding applied to the data  $\vsection = \vchannel \circ \dsection$. For the scatter plot, the mappings for the visual fiber components $\vfiber=(x,y, facecolors, s)$ are defined as
\begin{minted}{python}
cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
              'y': {'name': 'v2', 'encoder': lambda x: x},
              'facecolors': {'name':'v3', 'encoder': cmap}, 
              's':{'name': None , 'encoder': lambda _: itertools.repeat(.02)}}
\end{minted}
where the position \textit{(x,y)} \vchannel\ transformers are identity functions. The size \textit{s} transformer is not acting on a component of \dfiber, instead it is a \vchannel\ that returns a constant value. While size could be embedded inside the \mintinline{python}{assembly} function, it is added to the transformers to illustrate user configured visual parameters that could either be constant or mapped to a component in \dfiber. The identity and constant \vchannel\ are explicitly implemented here to demonstrate their implicit role in the visual pipeline, but they are somewhat optimized away in \mintinline{python}{Bar}. More complex encoders can be implemented as callable classes, such as 
\begin{minted}{python}
class Categorical:
    def __init__(self, mapping):
        # check that the conversion is to valid colors
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

@@

where \mintinline{python}{__init__} can validate that the output of the \vchannel\ is a valid element of the \vfiber\ component the \vchannel\ function is targeting. Creating a callable class also provides a simple way to swap out the specific (data, value) mapping without having to reimplement the validation or conversion logic. A test for equivariance can be implemented trivially
\begin{minted}{python}
def test_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    assert sorted(m1) == sorted(m2)
\end{minted}
but is currently factored out of the artist for clarity. In this example, \mintinline{python}{is_nominal} checks for equivariance of permutation group actions by applying the encoder to a set of values, shuffling values, and checking that (value, encoding) pairs remain the same.

@@

    def view(self, axes):
        table = defaultdict(list)
        for k in self.keys:
            table['index'] = k
            # on each iteration, add one (name, value) pair per component
            for (name, value) in zip(self.FB.fiber.keys(), self.tau(k)[1]):
                table[name].append(value)
        return table
\end{minted}
The view method returns a dictionary where the key is a fiber component name and the value is a list of values in the fiber component. The table is built one call to the section method \mintinline{python}{tau} at a time, guaranteeing that all the fiber component values are over the same \dbasepoint. Table has a \mintinline{python}{get} method as it is a method on Python dictionaries. In contrast, the line in \mintinline{python}{EdgeSimplex} is defined as the functions \mintinline{python}{_color,_xy} on each edge.

@@

    def view(self, axes):
        table = defaultdict(list)
        for k in self.keys:
            table['index'].append(k)
            # (name, value) pair, value is [x0, ..., xn] for x, y
            for (name, value) in zip(self.FB.fiber.keys(), self.tau(k, simplex)[1]):
                table[name].append(value)
\end{minted}
Unlike scatter, the line section method \mintinline{python}{tau} returns the functions on the edge evaluated on the interval [0,1]. By default these means each \mintinline{python}{tau} returns a list of 1000 x and y points and the associated color. As with scatter, \mintinline{python}{view} builds a table by calling \mintinline{python}{tau} for each \dbasepoint\. Unlike scatter, the line table is a list where each item contains a list of points. This bookkeeping of which data is on an edge is used by the \mintinline{python}{assembly} functions to bind segments to their visual properties.

@@

\subsection{Case Study: Penguins}
For this case study, we use the Palmer Penguins dataset\cite{gormanEcologicalSexualDimorphism2014, horstPalmerpenguinsPalmerArchipelago2020} since it is multivariate and has a varying number of penguins. We use a version of the data packaged as a pandas dataframe\cite{nakhaeeMcnakhaeePalmerpenguins2021} since that is a very commonly used Python labeled data structure. The wrapper is very thin because there is explicitly only one section.
\begin{minted}{python}
class DataFrame:
    def __init__(self, dataframe):
        self.FB = FiberBundle(K = {'tables':['vertex']},
                              F = dict(dataframe.dtypes))
        self._tau = dataframe.iloc
        self._view = dataframe

    def view(self, axes=None):
        return self._view
\end{minted}
Since the aim for this wrapper is to be very generic, here the fiber is set by querying the dataframe for its metadata. The \mintinline{python}{dtypes} are a list of column names and the datatype of the values in each column; this is the minimal amount of information the model requires to verify constraints. The pandas indexer is a key valued set of discrete vertices, so there is no need to repackage for the data interface.

@@

The stacked and grouped bar charts in figure~\ref{fig:code_bar_multi} are both out of \mintinline{python}{Bar} artists such that the difference between \mintinline{Python}{StackedBar} and \mintinline{Python}{GroupedBar} is specific to the ways in which the \mintinline{python}{Bar} are arranged differently. These two artists have identical \mintinline{{python}{__init__} and \mintinline{python}{draw} methods. As with \mintinline{python}{Bar}, the orientation is set on instantiation. In both these artists, we separate the transforms applied to only one component \mintinline{python}{transforms} and the case \mintinline{python}{mtransforms} where the same transform is applied to multiple components such that \votal\ has multiple components that map to the same retinal variable. 
\begin{minted}{python}
class StackedBar(martist.Artist):
    def __init__(self, data, transforms, mtransforms, orientation='v', *args, **kwargs):
        """
        Parameters
        ----------

        orientation: str, optional
            vertical: bars aligned along x axis, heights on y
            horizontal: bars aligned along y axis, heights on x   
        """
        super().__init__(*args, **kwargs)
        self.data = data
        self.orientation = orientation
        self.transforms = copy.deepcopy(transforms)
        self.mtransforms = copy.deepcopy(mtransforms)

    def assemble(self):
        view = self.data.view(self.axes)
        self.children = [] # list of bars to be rendered
        floor = 0
        for group in self.mtransforms:
            # pull out the specific group transforms
            group['floor'] = floor
            group.update(self.transforms)
            bar = Bar(self.data, group, self.orientation, transform=self.axes.transData)
            self.children.append(bar)
            floor += view[group['length']['name']]

    def draw(self, renderer, *args, **kwargs):
        # all the visual conversion gets pushed to child artists
        self.assemble()
        #self._transform = self.children[0].get_transform()
        for artist in self.children:
            artist.draw(renderer, *args, **kwargs)

\end{minted}
Since all the visual transformation is passed through to \mintinline{python}{Bar}, the \mintinline{python}{draw} method does not do any visual transformations. In \mintinline{Python}{StackedBar} the \mintinline{python}{view} is used to adjust the \mintinline{python}{floor} for every subsequent bar chart since a stacked bar chart is bar chart area marks concatenated together in the length parameter. In contrast, \mintinline{Python}{GroupedBar} does not even need the view, but instead keeps track of the relative position of each group of bars in the visual only variable \mintinline{python}{offset}.

\begin{minted}{python}
class GroupedBar(martist.Artist):
    def assemble(self):
        self.children = [] # list of bars to be rendered
        ngroups = len(self.mtransforms)

        for gid, group in enumerate(self.mtransforms):
            group.update(self.transforms)
            width = group.get('width', .8)
            group['width'] = width/ngroups
            group['offset'] = gid/ngroups*width 
            bar = Bar(self.data, group, self.orientation, transform=self.axes.transData)     
            self.children.append(bar)
\end{minted}{python}

Since the only difference between these two glyphs is in the composition of \mintinline{python}{Bar}, they take in the exact same transform specification dictionaries. \mintinline{python}{transform} dictates the position of the group, in this case by island the penguins are found on.

\begin{minted}{python}
transforms = {'position': {'name':'island',
              'encoder': position.Nominal({'Biscoe':0.1, 'Dream':1.1, 'Torgersen':2.1})}} 
group_transforms =  [{'length': {'name':'Adelie'},
                      'facecolors': {'name':"Adelie_s", 'encoder':cmap}},
                     {'length': {'name':'Chinstrap'},
                      'facecolors': {'name':"Chinstrap_s", 'encoder':cmap}}, 
                     {'length': {'name':'Gentoo'},
                      'facecolors': {'name':"Gentoo_s", 'encoder':cmap}}]
\end{minted}
\mintinline{python}{group_transforms} describes the group, and takes a list of dictionaries where each dictionary is the aesthetics of each group. That \mintinline{python}{position} and \mintinline{python}{length} are required parameters is enforced in the creation of the \mintinline{python}{Bar} artist. These means that these two artists have identical function signatures

\begin{minted}{python}
artistSB = bar.StackedBar(bt, ts, group_transforms)
artistGB = bar.GroupedBar(bt, ts, group_transforms)
\end{minted}

but differ in assembly \vmarkd. By decomposing the architecture into data, visual encoding, and assembly steps, we are able to build components that are more flexible and also more self contained than the existing code base. While very rough, this API demonstrates that the ideas presented in the math framework are implementable. In choosing a functional approach, if not implementation, we provide a framework for library developers to build reusable encoder \vchannel\, assembly \vmarkd and artists \vartist. We argue that  if these functions are built such that they are equivariant with respect to monoid actions and the graphic topology is a deformation retraction of the data topology, then the artist by definition will be a structure and property preserving map from data to graphic. 
\end{document}
